Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools

### 2024-03-27 paper.qmd

---
title: "Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools"
# subtitle: A case study of Ireland
bibliography: references.bib
csl: apa.csl
editor:
  markdown:
    wrap: sentence
execute: 
  echo: false
  message: false
number-sections: true
latex-clean: false
pdf-engine: pdflatex
format: 
  html: default
  # docx: default
# # # Uncomment for formatted version
  # elsevier-pdf:
  #   linestretch: 2
  #   keep-tex: true
  #   journal:
  #     name: European Transport Research Review
  #     formatting: preprint
  #     model: 3p
  #     cite-style: super
# # # #   # pdf: default

# # # Uncomment the following line to produce the cover page
# author:
#   - name: Robin Lovelace
#     email: r.lovelace@leeds.ac.uk
#     corresponding: true
#     affiliations:
#       - name: University of Leeds
#         address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
#   - name: Joey Talbot
#     affiliations:
#       - name: University of Leeds
#         address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
#   - name: Eugeni Vidal-Tortosa
#     affiliations:
#       - name: University of Leeds
#         address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
#   - name: Hussein Mahfouz
#     affiliations:
#       - name: University of Leeds
#         address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
#   - name: Elaine Brick
#     affiliations:
#       - name: AECOM
#         address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
#   - name: Peter Wright
#     affiliations:
#       - name: AECOM
#         address: Winslade House, Winslade Park, Manor Drive, Clyst St Mary, Exeter, EX5 1FY, UK
#   - name: "Gary O'Toole"
#     affiliations:
#       - name: AECOM
#         address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
#   - name: Dan Brennan
#     affiliations:
#       - name: Transport Infrastructure Ireland
#         address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
#   - name: Suzanne Meade
#     affiliations:
#       - name: Transport Infrastructure Ireland
#         address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
abstract: |
  This paper describes an approach for developing strategic cycle network planning tools. Based on our experience developing and deploying the Cycle Route Uptake and Scenario Estimation (CRUSE) Tool for Ireland, we outline the underlying methods, including disaggregation of origin-destination data, incorporation of additional trip purposes, routing, scenario generation, and development of an intuitive user interface that has been tested and used by practitioners. Commissioned and used by Transport Infrastructure Ireland, CRUSE provides estimates of current and potential future cycling levels under 'snapshot' scenarios to inform investment decisions. The publicly available results at https://cruse.bike/ enable planners, engineers, and other stakeholders to make more evidence-based decisions. CRUSE goes beyond previous work by: modeling networks at high spatial resolution; simulating multiple trip purposes (social, shopping, personal utility, recreational, and cycle touring) in addition to using official origin-destination data; and providing estimates of 'quietness' (a proxy for cyclist comfort and route preference) at the route segment level. Three network types --- 'Fastest', 'Balanced', and 'Quietest' --- help plan both arterial and residential cycle networks. Workshops and unstructured interviews with stakeholders show that the tool is already being used in practice to inform urban, inter-urban, and rural network planning. The approach is flexible and open-source, allowing the underlying ideas and code to be adapted, supporting effective cycling policies and interventions internationally.
keywords: 
  - Cycling
  - Open-source
  - Road Safety
  - Active Travel
  - Transport Planning
  - Collaborative Planning
date: last-modified
---

@@

::: {#fig-dublin}
<!-- ![](paste-1.png) -->

![](paste-8.png) <!-- ![](paste-3.png) -->
![](paste-9.png)

@@

<!-- ![](paste-5.png) -->

<!-- ![](paste-6.png) -->

@@

::: {#fig-route-types layout-nrow="2"}
![](naas_quietest_godutch.png){width="250"}

![](naas_fastest_godutch.png){width="250"}

![](newbridge_quietest_godutch.png){width="250"}

![](newbridge_fastest_godutch.png){width="250"}

@@

::: {#fig-scenarios}
<!-- ![](paste-7.png) -->

![](scenarios_kildare.png)

@@

::: {#fig-landing}
![](paste-10.png)

![](paste-11.png)

@@

}
")
# schematic
# htmlwidgets::saveWidget(schematic, "fig5.html")
# browseURL("fig5.html")
# webshot2::webshot(url = "fig5.html", file = "fig5.png")
# knitr::include_graphics("fig5.png")
schematic
```

```{=html}
<!-- ::: {#fig-user-stories}
![](fig5.png)

@@

```{r}
#| eval: false
#| echo: false
# Results for one city are shown below
# cities = c("dublin", "cork", "limerick", "galway")
library(tmap)
cities = c("dublin", "cork", "limerick")
city_name = "limerick"
city_buffer = zonebuilder::zb_zone(city_name)
city_buffer_5km = city_buffer |>
  filter(circle_id <= 2)
# qtm(city_buffer_5km)
city_buffer_union = sf::st_union(city_buffer_5km)
rnet_fastest = sf::read_sf(glue::glue("./../testout/{city_name}/rnet_combined_fastest.gpkg"))
rnet_fastest_city = sf::st_intersection(rnet_fastest, city_buffer_union)#
# nrow(rnet_fastest_city)
# table(rnet_fastest_city$`Bicycle (Go Dutch)`)
rnet_fastest_city_top = rnet_fastest_city |>
  filter(`Bicycle (Go Dutch)` != "1-10") |>
  filter(`Bicycle (Go Dutch)` != "0") |>
  mutate(across(`Bicycle (Baseline)`:`Quietness`, as.numeric)) |>
  slice_max(`Bicycle (Go Dutch)`, n = 10000)
m = tm_shape(rnet_fastest_city_top) +
  tm_lines(col = "Quietness", palette = "RdYlGn", lwd = 5)
tmap_save(m, "./rnet_fastest_city_top.png", width = 10, height = 10)
# Generalisation to balanced and quiestest route types:
route_types = c("fastest", "quietest")
dublin_rnets = purrr::map(route_types, function(route_type) {
  rnet = sf::read_sf(glue::glue("./../testout/{city_name}/rnet_combined_{route_type}.gpkg"))
  rnet_city = sf::st_intersection(rnet, city_buffer_union)
  rnet_city_top = rnet_city |>
    filter(`Bicycle (Go Dutch)` != "1-10") |>
    filter(`Bicycle (Go Dutch)` != "0") |>
    mutate(across(`Bicycle (Baseline)`:`Quietness`, as.numeric)) |>
    slice_max(`Bicycle (Go Dutch)`, n = 1000) |>
    mutate(`Route type` = route_type) |>
    rename(geometry = geom)
  rnet_city_top
})
dublin_rnets = do.call(rbind, dublin_rnets)
dublin_rnets$`Route type` = factor(dublin_rnets$`Route type`, levels = route_types)
m_types_dublin = tm_shape(dublin_rnets) +
  tm_lines(col = "col", lwd = "Bicycle (Go Dutch)", scale = 9) +
  tm_legend_hide() +
  tm_facets(by = "Route type", nrow = 1)

cities_rnet_types = purrr::map(cities, function(city_name) {
  city_buffer = zonebuilder::zb_zone(city_name)
  city_buffer_5km = city_buffer |>
    filter(circle_id <= 2)
  city_buffer_5km = stplanr::geo_bb(city_buffer_5km)
  city_buffer_union = sf::st_union(city_buffer_5km)
  route_types = c("fastest", "quietest")
  rnet_types = purrr::map(route_types, function(route_type) {
    rnet = sf::read_sf(glue::glue("./../testout/{city_name}/rnet_combined_{route_type}.gpkg"))
    rnet_city = sf::st_intersection(rnet, city_buffer_union)
    rnet_city_top = rnet_city |>
      filter(`Bicycle (Go Dutch)` != "1-10") |>
      filter(`Bicycle (Go Dutch)` != "0") |>
      mutate(across(`Bicycle (Baseline)`:`Quietness`, as.numeric)) |>
      # slice_max(`Bicycle (Go Dutch)`, n = 1000) |>
      # Title Case city name:
      mutate(`Route type` = str_to_title(route_type), `City` = str_to_title(city_name))
    rnet_city_top
  })
  rnet_types = do.call(rbind, rnet_types)
  rnet_types$`Route type` = factor(rnet_types$`Route type`, levels = str_to_title(route_types))
  rnet_types$`City` = factor(rnet_types$`City`, levels = str_to_title(cities))
  rnet_types
})
cities_rnet_types = do.call(rbind, cities_rnet_types)
# cities_rnet_types_1km = do.call(rbind, cities_rnet_types_1km)
summary(cities_rnet_types$`Bicycle (Go Dutch)`)
# Set max to 1000:
cities_rnet_types$`Bicycle (Go Dutch)` = pmin(cities_rnet_types$`Bicycle (Go Dutch)`, 1000)
m_types = tm_shape(cities_rnet_types) +
  tm_lines(col = "col", lwd = "Bicycle (Go Dutch)", scale = 20) +
  tm_legend_hide() +
  tm_facets_grid(rows = "City", columns = "Route type")
m_types
tmap_save(m_types, "./rnet_types.png", width = 12, height = 16)
browseURL("./rnet_types.png")
```

::: {#fig-city-results}
![](rnet_types.png)

>>>>

---
title: "Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools"
# subtitle: A case study of Ireland
bibliography: references.bib
csl: apa.csl
editor:
  markdown:
    wrap: sentence
execute: 
  echo: false
  message: false
number-sections: true
latex-clean: false
pdf-engine: pdflatex
knitr:
  opts_chunk:
    fig.path: images/
format: 
  # html: default
  # docx: default
# # # Uncomment for formatted version
  elsevier-pdf:
    linestretch: 2
    keep-tex: true
    journal:
      name: European Transport Research Review
      formatting: preprint
      model: 3p
      cite-style: super
# # # #   # pdf: default

# # Uncomment the following line to produce the cover page
author:
  - name: Robin Lovelace
    email: r.lovelace@leeds.ac.uk
    corresponding: true
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Joey Talbot
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Eugeni Vidal-Tortosa
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Hussein Mahfouz
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Elaine Brick
    affiliations:
      - name: AECOM
        address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
  - name: Peter Wright
    affiliations:
      - name: AECOM
        address: Winslade House, Winslade Park, Manor Drive, Clyst St Mary, Exeter, EX5 1FY, UK
  - name: "Gary O'Toole"
    affiliations:
      - name: AECOM
        address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
  - name: Dan Brennan
    affiliations:
      - name: Transport Infrastructure Ireland
        address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
  - name: Suzanne Meade
    affiliations:
      - name: Transport Infrastructure Ireland
        address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
abstract: |
  This paper describes an approach for developing strategic cycle network planning tools. Based on our experience developing and deploying the Cycle Route Uptake and Scenario Estimation (CRUSE) Tool for Ireland, we outline the underlying methods, including disaggregation of origin-destination data, incorporation of additional trip purposes, routing, scenario generation, and development of an intuitive user interface that has been tested and used by practitioners. Commissioned and used by Transport Infrastructure Ireland, CRUSE provides estimates of current and potential future cycling levels under 'snapshot' scenarios to inform investment decisions. The publicly available results at https://cruse.bike/ enable planners, engineers, and other stakeholders to make more evidence-based decisions. CRUSE goes beyond previous work by: modeling networks at high spatial resolution; simulating multiple trip purposes (social, shopping, personal utility, recreational, and cycle touring) in addition to using official origin-destination data; and providing estimates of 'quietness' (a proxy for cyclist comfort and route preference) at the route segment level. Three network types --- 'Fastest', 'Balanced', and 'Quietest' --- help plan both arterial and residential cycle networks. Workshops and unstructured interviews with stakeholders show that the tool is already being used in practice to inform urban, inter-urban, and rural network planning. The approach is flexible and open-source, allowing the underlying ideas and code to be adapted, supporting effective cycling policies and interventions internationally.
keywords: 
  - Cycling
  - Open-source
  - Road Safety
  - Active Travel
  - Transport Planning
  - Collaborative Planning
date: last-modified
---

@@

::: {#fig-dublin}
<!-- ![](images/paste-1.png) -->

![](images/paste-8.png) <!-- ![](images/paste-3.png) -->
![](images/paste-9.png)

@@

<!-- ![](images/paste-5.png) -->

<!-- ![](images/paste-6.png) -->

@@

::: {#fig-route-types layout-nrow="2"}
![](images/naas_quietest_godutch.png){width="250"}

![](images/naas_fastest_godutch.png){width="250"}

![](images/newbridge_quietest_godutch.png){width="250"}

![](images/newbridge_fastest_godutch.png){width="250"}

@@

::: {#fig-scenarios}
<!-- ![](images/paste-7.png) -->

![](images/scenarios_kildare.png)

@@

::: {#fig-landing}
![](images/paste-10.png)

![](images/paste-11.png)

@@

}
")
# schematic
# htmlwidgets::saveWidget(schematic, "images/fig5.html")
# browseURL("images/fig5.html")
# webshot2::webshot(url = "images/fig5.html", file = "images/fig5.png")
# knitr::include_graphics("images/fig5.png")
schematic
```

```{=html}
<!-- ::: {#fig-user-stories}
![](images/fig5.png)

@@

```{r}
#| eval: false
#| echo: false
# Results for one city are shown below
# cities = c("dublin", "cork", "limerick", "galway")
library(tmap)
cities = c("dublin", "cork", "limerick")
city_name = "limerick"
city_buffer = zonebuilder::zb_zone(city_name)
city_buffer_5km = city_buffer |>
  filter(circle_id <= 2)
# qtm(city_buffer_5km)
city_buffer_union = sf::st_union(city_buffer_5km)
rnet_fastest = sf::read_sf(glue::glue("./../testout/{city_name}/rnet_combined_fastest.gpkg"))
rnet_fastest_city = sf::st_intersection(rnet_fastest, city_buffer_union)#
# nrow(rnet_fastest_city)
# table(rnet_fastest_city$`Bicycle (Go Dutch)`)
rnet_fastest_city_top = rnet_fastest_city |>
  filter(`Bicycle (Go Dutch)` != "1-10") |>
  filter(`Bicycle (Go Dutch)` != "0") |>
  mutate(across(`Bicycle (Baseline)`:`Quietness`, as.numeric)) |>
  slice_max(`Bicycle (Go Dutch)`, n = 10000)
m = tm_shape(rnet_fastest_city_top) +
  tm_lines(col = "Quietness", palette = "RdYlGn", lwd = 5)
tmap_save(m, "./images/rnet_fastest_city_top.png", width = 10, height = 10)
# Generalisation to balanced and quiestest route types:
route_types = c("fastest", "quietest")
dublin_rnets = purrr::map(route_types, function(route_type) {
  rnet = sf::read_sf(glue::glue("./../testout/{city_name}/rnet_combined_{route_type}.gpkg"))
  rnet_city = sf::st_intersection(rnet, city_buffer_union)
  rnet_city_top = rnet_city |>
    filter(`Bicycle (Go Dutch)` != "1-10") |>
    filter(`Bicycle (Go Dutch)` != "0") |>
    mutate(across(`Bicycle (Baseline)`:`Quietness`, as.numeric)) |>
    slice_max(`Bicycle (Go Dutch)`, n = 1000) |>
    mutate(`Route type` = route_type) |>
    rename(geometry = geom)
  rnet_city_top
})
dublin_rnets = do.call(rbind, dublin_rnets)
dublin_rnets$`Route type` = factor(dublin_rnets$`Route type`, levels = route_types)
m_types_dublin = tm_shape(dublin_rnets) +
  tm_lines(col = "col", lwd = "Bicycle (Go Dutch)", scale = 9) +
  tm_legend_hide() +
  tm_facets(by = "Route type", nrow = 1)

cities_rnet_types = purrr::map(cities, function(city_name) {
  city_buffer = zonebuilder::zb_zone(city_name)
  city_buffer_5km = city_buffer |>
    filter(circle_id <= 2)
  city_buffer_5km = stplanr::geo_bb(city_buffer_5km)
  city_buffer_union = sf::st_union(city_buffer_5km)
  route_types = c("fastest", "quietest")
  rnet_types = purrr::map(route_types, function(route_type) {
    rnet = sf::read_sf(glue::glue("./../testout/{city_name}/rnet_combined_{route_type}.gpkg"))
    rnet_city = sf::st_intersection(rnet, city_buffer_union)
    rnet_city_top = rnet_city |>
      filter(`Bicycle (Go Dutch)` != "1-10") |>
      filter(`Bicycle (Go Dutch)` != "0") |>
      mutate(across(`Bicycle (Baseline)`:`Quietness`, as.numeric)) |>
      # slice_max(`Bicycle (Go Dutch)`, n = 1000) |>
      # Title Case city name:
      mutate(`Route type` = str_to_title(route_type), `City` = str_to_title(city_name))
    rnet_city_top
  })
  rnet_types = do.call(rbind, rnet_types)
  rnet_types$`Route type` = factor(rnet_types$`Route type`, levels = str_to_title(route_types))
  rnet_types$`City` = factor(rnet_types$`City`, levels = str_to_title(cities))
  rnet_types
})
cities_rnet_types = do.call(rbind, cities_rnet_types)
# cities_rnet_types_1km = do.call(rbind, cities_rnet_types_1km)
summary(cities_rnet_types$`Bicycle (Go Dutch)`)
# Set max to 1000:
cities_rnet_types$`Bicycle (Go Dutch)` = pmin(cities_rnet_types$`Bicycle (Go Dutch)`, 1000)
m_types = tm_shape(cities_rnet_types) +
  tm_lines(col = "col", lwd = "Bicycle (Go Dutch)", scale = 20) +
  tm_legend_hide() +
  tm_facets_grid(rows = "City", columns = "Route type")
m_types
tmap_save(m_types, "./images/rnet_types.png", width = 12, height = 16)
browseURL("./images/rnet_types.png")
```

::: {#fig-city-results}
![](images/rnet_types.png)

### 2024-03-27 paper.tex

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools},
  pdfkeywords={Cycling, Open-source, Road Safety, Active
Travel, Transport Planning, Collaborative Planning},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{Cycle Route Uptake and Scenario Estimation (CRUSE): an approach
for developing strategic cycle network planning tools}

@@

\includegraphics{paste-8.png} \includegraphics{paste-9.png}

@@

\begin{minipage}{0.50\linewidth}
\includegraphics[width=2.60417in,height=\textheight]{naas_quietest_godutch.png}\end{minipage}%
%
\begin{minipage}{0.50\linewidth}
\includegraphics[width=2.60417in,height=\textheight]{naas_fastest_godutch.png}\end{minipage}%
\newline
\begin{minipage}{0.50\linewidth}
\includegraphics[width=2.60417in,height=\textheight]{newbridge_quietest_godutch.png}\end{minipage}%
%
\begin{minipage}{0.50\linewidth}
\includegraphics[width=2.60417in,height=\textheight]{newbridge_fastest_godutch.png}\end{minipage}%

@@

\includegraphics{scenarios_kildare.png}

@@

\includegraphics{paste-10.png}

\includegraphics{paste-11.png}

@@

\includegraphics{fig-user-stories-1.pdf}

@@

\includegraphics{rnet_types.png}

>>>>

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools},
  pdfauthor={Robin Lovelace; Joey Talbot; Eugeni Vidal-Tortosa; Hussein Mahfouz; Elaine Brick; Peter Wright; Gary O'Toole; Dan Brennan; Suzanne Meade},
  pdfkeywords={Cycling, Open-source, Road Safety, Active
Travel, Transport Planning, Collaborative Planning},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{Cycle Route Uptake and Scenario Estimation (CRUSE): an approach
for developing strategic cycle network planning tools}
\author[1]{Robin Lovelace%
\corref{cor1}%
}
 \ead{r.lovelace@leeds.ac.uk} 
\author[1]{Joey Talbot%
%
}

\author[1]{Eugeni Vidal-Tortosa%
%
}

\author[1]{Hussein Mahfouz%
%
}

\author[2]{Elaine Brick%
%
}

\author[3]{Peter Wright%
%
}

\author[2]{Gary O'Toole%
%
}

\author[4]{Dan Brennan%
%
}

\author[4]{Suzanne Meade%
%
}

\affiliation[1]{organization={University of
Leeds},addressline={Institute for Transport Studies, Leeds, LS2 9JT,
United Kingdom},postcodesep={}}
\affiliation[2]{organization={AECOM},addressline={Unit 6, Galway
Technology Park, Parkmore, Galway, Ireland},postcodesep={}}
\affiliation[3]{organization={AECOM},addressline={Winslade House,
Winslade Park, Manor Drive, Clyst St Mary, Exeter, EX5 1FY,
UK},postcodesep={}}
\affiliation[4]{organization={Transport Infrastructure
Ireland},addressline={Parkgate Business Centre, Parkgate Street, Dublin
8, D08 DK10, Ireland},postcodesep={}}

@@

\includegraphics{images/paste-8.png}
\includegraphics{images/paste-9.png}

@@

\begin{minipage}{0.50\linewidth}
\includegraphics[width=2.60417in,height=\textheight]{images/naas_quietest_godutch.png}\end{minipage}%
%
\begin{minipage}{0.50\linewidth}
\includegraphics[width=2.60417in,height=\textheight]{images/naas_fastest_godutch.png}\end{minipage}%
\newline
\begin{minipage}{0.50\linewidth}
\includegraphics[width=2.60417in,height=\textheight]{images/newbridge_quietest_godutch.png}\end{minipage}%
%
\begin{minipage}{0.50\linewidth}
\includegraphics[width=2.60417in,height=\textheight]{images/newbridge_fastest_godutch.png}\end{minipage}%

@@

\includegraphics{images/scenarios_kildare.png}

@@

\includegraphics{images/paste-10.png}

\includegraphics{images/paste-11.png}

@@

\includegraphics{images/fig-user-stories-1.pdf}

@@

\includegraphics{images/rnet_types.png}
